---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Olivia Zhang"
date: "02/14/2025"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
#install.packages("devtools")
#devtools::install_github("rstudio-education/dsbox")

library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

The Denny's dataset has `r nrow(dennys)` rows and `r ncol(dennys)` columns/variables. Each row represent a Denny's location, and the variables are address, city, state, zip, longitude, and latitude.

### Exercise 2

The La Quinta's dataset has `r nrow(laquinta)` rows and `r ncol(laquinta)` columns/variables. Each row represent a La Quinta's location, and the variables are address, city, state, zip, longitude, and latitude.

### Exercise 3

There are some La Quintas' outside of US. They are in countries like Canada, Mexico, 
China, New Zealand, Turkey, United Arab Emirates, Chile, Colombia, Ecuador. However, 
I don't think there's any Danny's outside of the US. 

### Exercise 4

There are two ways I can think of to determine whether there's any establishments that are not in the US. 
1. We can create a new variable called US and check weather the state variable in two datasets 
are the same or different from the abbreviation variable from the states dataset. 
2. If we know the longitude and latitude of the US border, techniquely, we can filter out the 
establishments outside of US. 

### Exercise 5

```{r filter-dn}
dn <- dennys %>%
  filter(!(state %in% states$abbreviation))
dn
```
There are `r nrow(dn)` Denny's locations outside the US. 

### Exercise 6

```{r dn-country}
dennys <- dennys %>%
  mutate(country = "United States")
dennys
```

### Exercise 7

```{r filter-lq}
lq <- laquinta %>%
  filter(!(state %in% states$abbreviation))
lq
```
There are `r nrow(lq)` Denny's locations outside the US. 

### Exercise 8
 
```{r lq-country}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT" ~ "Colombia",
    state == "FM" ~ "Honduras",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL") ~ "Mexico",
  ))
```

```{r lq-filter}
laquinta <- laquinta %>%
  filter(country == "United States")
```

### Exercise 9

```{r count}
dennys %>%
  count(state) %>%
  arrange(desc(n))

laquinta %>%
  count(state) %>%
  arrange(desc(n))
```
California (n = 403) has the most Denny's locations and Delaware has the least (n = 1); Texas (n = 237) has the most La Quinta's locations and Maine has the least (n = 1). Due to my limited knowledge of the US demographic, I guess it's not surprising, since Texas is near Mexico and Maine is far north. I'm not sure why the location pattern of Danny's is this way though.

```{r join}
dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))

laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```

### Exercise 10

```{r cal-ptsm}
dennys <- dennys %>%
  mutate(per_th_sq_mile = )
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")
```


```{r set-id}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")
```

```{r bind}
dn_lq <- bind_rows(dennys, laquinta)
```

```{r plot}
ggplot(dn_lq, mapping = aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point()
```

### Exercise 11

```{r plot-nc}
dn_lq %>%
  filter(state == "NC") %>%
  ggplot(dn_lq, mapping = aes(
    x = longitude,
    y = latitude,
    color = establishment
    )) +
  geom_point(alpha = .7) +
  labs(
  title = "Danny's and La Quinta Locations",
  subtitle = "in North Carolina",
  x = "Longitude of establishements", 
  y = "Latitude of establishements", 
  color = "Establishment"
     )
```
Mitch Hedberg’s joke mostly hold in North Carolina except for a few La Quintas on the upper left corner of the plot, which are distant from their nearest Denny's. 

### Exercise 12

```{r plot-tx}
dn_lq %>%
  filter(state == "TX") %>%
  ggplot(dn_lq, mapping = aes(
    x = longitude,
    y = latitude,
    color = establishment
    )) +
  geom_point(alpha = .5) +
  labs(
  title = "Danny's and La Quinta Locations",
  subtitle = "in Texas",
  x = "Longitude of establishements", 
  y = "Latitude of establishements", 
  color = "Establishment"
     )
```
Mitch Hedberg’s joke seems to hold less in Texas, since there are more La Quintas than Denny's, and not all La Quinta has a Denny's next to it. However, all Denny's are near La Quintas.
